---
title: "Code Highlighting with demoR"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{code_highlighting}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
# knitr::opts_chunk$set(
#   collapse = TRUE,
#   comment = "#>"
# )
```


```{r setup, message = FALSE, warning = FALSE}
library(demoR)
library(dplyr)
library(ggplot2)
library(stringr)
```

## Introduction

The primary goal of the `demoR` package is to simplify the presentation of R code.

It is common to show source code, in addition to code output, as part of a conference talk, workshop, or lecture.  Often, we want to call attention to certain aspects of the code. 

For example, suppose you want to illustrate to a new learner the use of the pipe `%>%`.  You might want to create a slide that shows the following:

```{r, echo = FALSE}

demo_code('

iris %>%
  group_by(Species) %>%
  summarize(mean(Sepal.Length))

') %>%
  hlt_regexp(fixed("%>%"))


```


Without `demoR`, your approach might be to type your code into your code chunk, copy-paste it to a string, and manually format that string using html.  What a headache!  

In this vignette, we will show you how to use the `demoR` package to simultaneously run and format code.

## demo_code Objects

The core element of the `demoR` package is an object of the class `demo_code`.  These objects contain both the natural output of your R code and information to let `knitr` know how display the source code.


### Demoing from a code chunk

The cleanest way to make a `demo_code` object is by referencing a named code chunk in R Markdown.

```{r how_to_pipe, include = FALSE}

iris %>%
  group_by(Species) %>%
  summarize(mean(Sepal.Length))

```

For example:

````

```{r how_to_pipe, include = FALSE}`r ''`

iris %>%
  group_by(Species) %>%
  summarize(mean(Sepal.Length))

```

````


You can use the chunk label to automatically create a `demo_code` object using `create_demo()`.  Here we pass the object to `hlt_fixed()` to show our source code with the pipe operators highlighted in yellow.

````

```{r, echo = FALSE}`r ''`

demo_chunk("how_to_pipe") %>% hlt_fixed("%>%")
  
```

````


Note that the `create_demo` step should be in a separate chunk from the code you are demoing, since it is not itself part of the source code you wish to display.

With the above two code chunks in our source file, the resulting knitted output looks like this:

```{r, echo = FALSE}
demo_chunk("how_to_pipe") %>% 
  hlt_fixed("%>%") 
```

### Demoing from a code string

You can also use the `demo_code()` function to create a `demo_code` object directly from a string of R code. For example

```{r, eval = FALSE}
demo_code('

iris %>%
  group_by(Species) %>%
  summarize(mean(Sepal.Length))

') %>%
  hlt_fixed("%>%")
```

produces

```{r, echo = FALSE}
demo_code('

iris %>%
  group_by(Species) %>%
  summarize(mean(Sepal.Length))

') %>%
  hlt_fixed("%>%")
```


This option for `demo_code` creation is particularly nice if you want to show "bad" code that cannot normally be evaluated in a chunk.  For example:

```{r, error = TRUE}
demo_code('

mean(1:10

') %>%
  hlt_fixed("(")
```



## The `hlt_*` functions

The advantage of a `demo_code` object is that you can add formatting to the source code without altering the output, through the suite of highlightint (`hlt`) functions

### hlt_regexp

The main function you will use is `hlt_regexp()`.  This takes as arguments:

* A string or a `demo_code` object

* A regular expression

* Any number of formatting parameters

If no formatting parameters are supplied, `hlt_*` will default to ordinary yellow-background highlighting.

`hlt_regexp` returns a `demo_code` object, so it is pipe friendly!

Suppose you want to highlight the pipe (`%>%`) in yellow, highlight the variable name `Sepal.Length` in pink, and change the text color of `Species` to blue

Your code chunk:

```{r, eval = FALSE}
demo_code('

iris %>%
  group_by(Species) %>%
  summarize(mean(Sepal.Length))

') %>%
  hlt_regexp(fixed("%>%")) %>%
  hlt_regexp("Sepal.Length", background = "pink") %>%
  hlt_regexp("Species", color = "CornflowerBlue")
```

Your knitted output:

```{r, echo = FALSE}
demo_code('

iris %>%
  group_by(Species) %>%
  summarize(mean(Sepal.Length))

') %>%
  hlt_regexp(fixed("%>%")) %>%
  hlt_regexp("Sepal.Length", background = "Pink") %>%
  hlt_regexp("Species", color = "CornflowerBlue")
```


### Syntax highlighting

`demoR` also includes a few shortcuts for highlighting specific aspsects of R source code.  Currently, these functions are:

* `hlt_funs()` for *functions*

* `hlt_args()` for *arguments to functions*

* `hlt_input_vals()` for *values assigned to function arguments*

For example:

```{r, fig.width = 7, fig.height = 5}
demo_code('

ggplot(iris, aes(x = Sepal.Length, fill = Species)) +
  geom_histogram(alpha = 0.5)

') %>%
  hlt_args(color = "CornflowerBlue") %>%
  hlt_funs(color = "Coral", underline = TRUE) %>%
  hlt_input_vals(background = "Aquamarine") %>%
  hlt_regexp("Sepal.Length", background = "pink")
```


## Errata 

### Defining objects

One nice feature of a `demo_code` object, as compared to a string, it that it evaluates when run.  This means that you can define objects in your demo code, and use them later in your analysis as you normally would:


```{r}

demo_code('foo <- mean(1:10)') %>% hlt_funs()

foo + 5
```


A word of caution: Make sure you define your objects in your code string, not outside the `demo_code()` function!  For example, the following approach has two problems:

1. `foo` contains the `demo_code` object, not the result of `mean(1:10)`, so  `foo + 5` throws an error.

2. The `demo_code` object is stored as `foo` rather than printed, so no highlighted code is included in the knitted output.

```{r, error = TRUE}
foo <- demo_code('mean(1:10)') %>% hlt_funs()

foo + 5
```


### A note about colors

`demoR` gives you complete freedom to choose the colors of your highlighted elements, so long as the color name is a [recognized html name](https://www.w3schools.com/colors/colors_names.asp) or a hex code.

However, please remember to be judicious in your color choices, and to keep in mind [how your colors appear to colorblind individuals](https://venngage.com/blog/color-blind-friendly-palette/).

